---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Data Sources

In this project, we used several data sources. 

## Gun Violence Data

Gun violence data records each gun violence incidents.  

We got this dataset from [kaggle](https://www.kaggle.com/jameslko/gun-violence-data), the original dataset is from [gunviolencearchive.org](https://www.gunviolencearchive.org/). Since the data from GVA limits the number of incidents that are returned from a single query, and the website's "Export to CSV" functionality missed some crucial fields, it's not ready to be used. So, we used the data from kaggle, these data were obtained using web scraping. And we obtained the meaning of each column from GVA.  

This size of this dataset is 239,677 $\times$ 29. This dataset has 29 variables, and each row represents a record of gun violence.
```{r include=FALSE}
library(readr)
data <- read.csv("dataset/clean gun-violence-data_2013-3-2018-3_new.csv")
```

**Desctiption of each columns**  
```{r echo=FALSE}
res=data.frame(Colnames=c('n_killed','n_injured','Children/Teens Involve','Mass Shooting','Home Invasion','Officer Involved Shooting','longitude','latitude'),Meaning=c('Total number of killed people in the shooting','Total number of injured people in the shooting','Whether the incident involved children or teenagers','Whether the incident is mass shooting','Whether the incident is home invasion', 'Whether the incident involved officer','Longitude of gun violence incident','Latitude of gun violence incident'))
res=knitr::kable(res)
res
```

We used variables like "date", "state", "n_killed", "n_injured", "longitude" and "latitude". Because the original file is quite large ( >150MB ), and some columns cannot be directly used, therefore, we only selected some characters of each incident and built up a new dataset.  

## FBI NICS Firearm Gun Background Check Data

Background Check Data is used as the estimate of gun sales in United States.   

This dataset is about gun background check, and the content of background checking is to check whether the transfer of firearms is illegal. We obtained this data from [github](https://github.com/BuzzFeedNews/nics-firearm-background-checks/blob/master/README.md), it says that "The FBI’s NICS numbers are widely accepted as the best proxy for total gun sales in a given time period". The original dataset is from [National Instant Criminal Background Check System](https://www.fbi.gov/services/cjis/nics). Since the original dataset from FBI is in PDF form, we use the data from github.

The size of this data is 13860 $\times$ 29. This dataset has 29 variables. And each row represents a record of the firearm transfer. There are three types of firearms in the dataset, handgun, long gun and others.  
```{r include=FALSE}
data <- read.csv("dataset/nics-firearm-background-checks copy.csv")
```

**Desctiption of each columns**  
```{r echo=FALSE}
res=data.frame(Colnames=c('permit','prepawn_##',"redemption_##",'returned_##','rentals_##','private_sale_##','return_to_seller_##','totals','X28'),Meaning=c('Total number of transfers which are permitted','Total number of prepawn transfers','Total number of redemption transfers','Total number of transfers of returning firearm to the owner','Total number of rental transfers', 'Total number of private sale transfers','Total number of transfers of returning firearm to the seller' ,'Total number of incidents','The abbreviation of states'))
res=knitr::kable(res)
res
```

"##" means gun types: handgun, long gun and others. We are only interested in the total number of the gun background data "totals". 

## Population Data

The population data was obtained from [Census](https://www.census.gov/newsroom/press-kits/2018/pop-estimates-national-state.html). Since the number of gun violence and background check is based on population, we would like to focus on rate numbers rather than total numbers. We only got the population of each state from year 2010 to year 2018.  
```{r include=FALSE}
data <- read.csv("dataset/clean_population_by_state.csv")
```

**Desctiption of each columns**  
```{r echo=FALSE}
res=data.frame(Colnames=c('X1','state','year','population'),Meaning=c('Index of each column','State name','year','population of the year'))
res=knitr::kable(res)
res
```

## GDP Data  

We would also like to find whether gun violence is related to ecomonic data, so we used GDP of each state to reflect the economic situation. We obtained this dataset about GDP from [Bureau of Economic Analysis](https://www.bea.gov/data/gdp/gdp-state). 
Originally, each state has its own GDP dataset from 1999 to 2018. We combine those datasets and used total GDP as the representative of economic status.

```{r include=FALSE}
GDP<-read.csv("dataset/GDP_states.csv")
```

**Desctiption of each columns**  
```{r echo=FALSE}
res=data.frame(Colnames=c('GeoName','ComponentName',"Description",'Xyyyy'),Meaning=c('State name','Gross domestic product (GDP) by state Millions of current dollars ','total GDP','GDP of yyyy year'))
res=knitr::kable(res)
res
```