# Results

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(ggplot2)
library(tidyverse)
library(stringr)
library(GGally)
library(parcoords)
library(tidyr)
library(r2d3)
library(dplyr)
library(ggmap)
library(maps)
library(mapdata)
library(readr)
library(gridExtra)
library(plotly)
```

```{r include=FALSE}
# read file
gun_vio=read.csv("dataset/clean gun-violence-data_2013-3-2018-3_new.csv")
gun_vio["year"]=str_sub( as.character(gun_vio$date),1,4)
gun_vio<-gun_vio[which(as.numeric(as.character(gun_vio$year))>2013)  ,]
gun_vio$date<-as.Date(gun_vio$date)
gun_vio$state<-as.character(gun_vio$state)
gun_vio["year"]<-format(as.Date(gun_vio$date, format="%d/%m/%Y"),"%Y")
gun_vio["month"]<-format(as.Date(gun_vio$date, format="%d/%m/%Y"),"%m")

gun_vio_n <- read.csv("dataset/gun_violence_sum.csv")
back_check=read.csv("dataset/nics-firearm-background-checks copy.csv")
GDP<-read.csv("dataset/GDP_states.csv")
GDP_sub<-GDP %>% select(-c(1,2,4,5,6,7,8,9))

gun_character <- read.csv("dataset/gun-violence-character_year.csv")
population_tp=read.csv("dataset/clean_population_by_state.csv")
population<-select(population_tp,2,3,4)

back_check$month<-as.character(back_check$month)
back_check["year"]=str_sub(back_check$month,1,4)

states_party<-read.csv("dataset/states_party.csv")
```

## Exploring the Gun Violence Data

We start from an interesting question: Where is the most gun threatened state in the US? Then we will explore the gun violence data, hoping to find the answers from it. 

### How many gun violence incidents happened each year in each state?

We calculated the the sum of incidents number of each state from 2014 to 2018, the plots are shown below:

```{r include=FALSE}
#incident number
gun_vio_states<-gun_vio%>%
  group_by(state,year)%>%
  summarize(incident_number=n())
gun_vio_states<-as.data.frame(gun_vio_states)
gun_vio_states_2014 <- gun_vio_states %>%
  filter(year=="2014")
gun_vio_states_2015 <- gun_vio_states %>%
  filter(year=="2015")
gun_vio_states_2016 <- gun_vio_states %>%
  filter(year=="2016")
gun_vio_states_2017 <- gun_vio_states %>%
  filter(year=="2017")
gun_vio_states_2018 <- gun_vio_states %>%
  filter(year=="2018")

sum_2014 <- ggplot(gun_vio_states_2014 ,aes(x=fct_reorder(state,incident_number),y=incident_number,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Number of Incidents by States in 2014")+
  labs(y="incident number",x="States")+
  theme(legend.position = "none")

sum_2015 <- ggplot(gun_vio_states_2015 ,aes(x=fct_reorder(state,incident_number),y=incident_number,fill="#b30000"))+
  scale_fill_manual(values=c("#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Number of Incidents by States in 2015")+
  labs(y="incident number",x="States")+
  theme(legend.position = "none")

sum_2016 <- ggplot(gun_vio_states_2016 ,aes(x=fct_reorder(state,incident_number),y=incident_number,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Number of Incidents by States in 2016")+
  labs(y="incident number",x="States")+
  theme(legend.position = "none")

sum_2017 <- ggplot(gun_vio_states_2017 ,aes(x=fct_reorder(state,incident_number),y=incident_number,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Number of Incidents by States in 2017")+
  labs(y="incident number",x="States")+
  theme(legend.position = "none")

sum_2018 <- ggplot(gun_vio_states_2018 ,aes(x=fct_reorder(state,incident_number),y=incident_number,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Number of Incidents by States in 2018")+
  labs(y="incident number",x="States")+
  theme(legend.position = "none")
```

```{r echo=FALSE, fig.width=9,fig.height=16}
grid.arrange(sum_2014,sum_2015,sum_2016,sum_2017,sum_2018,ncol=2)
```

The bar in 2018 is shorter than the bars in other years, it's because we only have the record from Jan to March in 2018. We could find that the number of incidents is increasing by year.   
We also could find that in 2014 and 2018, California has the largest number of incidents; while in 2015, 2016 and 2017, Illinois has the largest number of incidents.  
But, does the large number comes from the large population? Maybe it's better to calculate the number of incidents per 100000 person. The plots are shown below:

```{r include=FALSE}
#incident number per 100000 person
gun_vio_adjust = merge(gun_vio_states,population, by.x=c("state", "year"), by.y=c("state", "year"))
gun_vio_adjust["incident_num_avg"]<-gun_vio_adjust$incident_number/gun_vio_adjust$population*100000
gun_vio_adjust<-select(gun_vio_adjust,1,2,5)

gun_vio_adjust_2014 <- gun_vio_adjust %>%
  filter(year=="2014")
gun_vio_adjust_2015 <- gun_vio_adjust %>%
  filter(year=="2015")
gun_vio_adjust_2016 <- gun_vio_adjust %>%
  filter(year=="2016")
gun_vio_adjust_2017 <- gun_vio_adjust %>%
  filter(year=="2017")
gun_vio_adjust_2018 <- gun_vio_adjust %>%
  filter(year=="2018")

avg_2014 <- ggplot(gun_vio_adjust_2014 ,aes(x=fct_reorder(state,incident_num_avg),y=incident_num_avg,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,170)+
  coord_flip()+
  ggtitle("Incident Number per 100k person in 2014")+
  labs(y="average incident number",x="States")+
  theme(legend.position = "none")

avg_2015 <- ggplot(gun_vio_adjust_2015 ,aes(x=fct_reorder(state,incident_num_avg),y=incident_num_avg,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,170)+
  coord_flip()+
  ggtitle("Incident Number per 100k person in 2015")+
  labs(y="average incident number",x="States")+
  theme(legend.position = "none")

avg_2016 <- ggplot(gun_vio_adjust_2016 ,aes(x=fct_reorder(state,incident_num_avg),y=incident_num_avg,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,170)+
  coord_flip()+
  ggtitle("Incident Number per 100k person in 2016")+
  labs(y="average incident number",x="States")+
  theme(legend.position = "none")

avg_2017 <- ggplot(gun_vio_adjust_2017 ,aes(x=fct_reorder(state,incident_num_avg),y=incident_num_avg,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,170)+
  coord_flip()+
  ggtitle("Incident Number per 100k person in 2017")+
  labs(y="average incident number",x="States")+
  theme(legend.position = "none")

avg_2018 <- ggplot(gun_vio_adjust_2018 ,aes(x=fct_reorder(state,incident_num_avg),y=incident_num_avg,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,170)+
  coord_flip()+
  ggtitle("Incident Number per 100k person in 2018")+
  labs(y="average incident number",x="States")+
  theme(legend.position = "none")
```

```{r echo=FALSE, fig.width=9,fig.height=16}
grid.arrange(avg_2014,avg_2015,avg_2016,avg_2017,avg_2018,ncol=2)
```

The resultes turns out very different, District of Columbia moves to the top of the bar from 2014 to 2018; Alaska becomes the second most in 2015 and 2016. It might because the population in these states are quite small. 

### How many casualties each year in each state?

The number of casualties in every incident is different,the plots of total number of casualties each year by state are shown below: 

```{r include=FALSE}

gun_vio_n_2014 <- gun_vio_n %>%filter(year=="2014")
gun_vio_n_2015 <- gun_vio_n %>%filter(year=="2015")
gun_vio_n_2016 <- gun_vio_n %>%filter(year=="2016")
gun_vio_n_2017 <- gun_vio_n %>%filter(year=="2017")
gun_vio_n_2018 <- gun_vio_n %>%filter(year=="2018")

sum_n_2014 <- ggplot(gun_vio_n_2014 ,aes(x=fct_reorder(state,total_TOLL),y=total_TOLL,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Number of Casualties in 2014")+
  labs(y="number of casualties",x="States")+
  theme(legend.position = "none")

sum_n_2015 <- ggplot(gun_vio_n_2015 ,aes(x=fct_reorder(state,total_TOLL),y=total_TOLL,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Number of Casualties in 2015")+
  labs(y="number of casualties",x="States")+
  theme(legend.position = "none")

sum_n_2016 <- ggplot(gun_vio_n_2016 ,aes(x=fct_reorder(state,total_TOLL),y=total_TOLL,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Number of Casualties in 2016")+
  labs(y="number of casualties",x="States")+
  theme(legend.position = "none")

sum_n_2017 <- ggplot(gun_vio_n_2017 ,aes(x=fct_reorder(state,total_TOLL),y=total_TOLL,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Number of Casualties in 2017")+
  labs(y="number of casualties",x="States")+
  theme(legend.position = "none")

sum_n_2018 <- ggplot(gun_vio_n_2018 ,aes(x=fct_reorder(state,total_TOLL),y=total_TOLL,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,5100)+
  coord_flip()+
  ggtitle("Number of Casualties in 2018")+
  labs(y="number of casualties",x="States")+
  theme(legend.position = "none")

```

```{r echo=FALSE, fig.width=9 ,fig.height=15}
grid.arrange(sum_n_2014,sum_n_2015,sum_n_2016,sum_n_2017,sum_n_2018,ncol=2)
```

The bar in 2018 is shorter than the bars in other years, it's because we only have the record from Jan to March in 2018. We could find this pattern is quite similar to the pattern of incident number. Illinois has the largest number from 2015 to 2018; California has the largest number in 2014. Then we would like to look at the number of casualties per 100k person. The plots are shown below:  

```{r include=FALSE}
gun_vio_n_adjust = merge(gun_vio_n,population, by.x=c("state", "year"), by.y=c("state", "year"))

gun_vio_n_adjust['sum_killed_avg']<-gun_vio_n_adjust$total_TOLL/gun_vio_n_adjust$population*100000
gun_vio_n_adjust<-select(gun_vio_n_adjust,1,2,6)

gun_vio_adjust_2014 <- gun_vio_n_adjust %>%
  filter(year=="2014")
gun_vio_adjust_2015 <- gun_vio_n_adjust %>%
  filter(year=="2015")
gun_vio_adjust_2016 <- gun_vio_n_adjust %>%
  filter(year=="2016")
gun_vio_adjust_2017 <- gun_vio_n_adjust %>%
  filter(year=="2017")
gun_vio_adjust_2018 <- gun_vio_n_adjust %>%
  filter(year=="2018")


avg_n_2014 <- ggplot(gun_vio_adjust_2014 ,aes(x=fct_reorder(state,sum_killed_avg),y=sum_killed_avg,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,70)+
  coord_flip()+
  ggtitle("Number of Casualties per 100k person in 2014")+
  labs(y="average number of casualties",x="States")+
  theme(legend.position = "none")

avg_n_2015 <- ggplot(gun_vio_adjust_2015 ,aes(x=fct_reorder(state,sum_killed_avg),y=sum_killed_avg,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,70)+
  coord_flip()+
  ggtitle("Number of Casualties per 100k person in 2015")+
  labs(y="average number of casualties",x="States")+
  theme(legend.position = "none")

avg_n_2016 <- ggplot(gun_vio_adjust_2016 ,aes(x=fct_reorder(state,sum_killed_avg),y=sum_killed_avg,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,70)+
  coord_flip()+
  ggtitle("Number of Casualties per 100k person in 2016")+
  labs(y="average number of casualties",x="States")+
  theme(legend.position = "none")

avg_n_2017 <- ggplot(gun_vio_adjust_2017 ,aes(x=fct_reorder(state,sum_killed_avg),y=sum_killed_avg,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,70)+
  coord_flip()+
  ggtitle("Number of Casualties per 100k person in 2017")+
  labs(y="average number of casualties",x="States")+
  theme(legend.position = "none")

avg_n_2018 <- ggplot(gun_vio_adjust_2018 ,aes(x=fct_reorder(state,sum_killed_avg),y=sum_killed_avg,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  ylim(0,70)+
  coord_flip()+
  ggtitle("Number of Casualties per 100000 person in 2018")+
  labs(y="average number of casualties",x="States")+
  theme(legend.position = "none")
```

```{r echo=FALSE, fig.width=9,fig.height=15}
grid.arrange(avg_n_2014,avg_n_2015,avg_n_2016,avg_n_2017,avg_n_2018,ncol=2)
```

This pattern is also similar to the pattern of incident number per 100k person, District of Columbia is at the top of the bar chart from 2014 to 2018. 

### Where did these gun violence incidents happen?

Then, we zoomed in the map to find the exact location of each gun incident. We ploted each event as a red dot on the map. The following plot shows the gun violence in California in 2017, 

```{r echo=FALSE}
names(gun_vio)[5] <- "lat"
names(gun_vio)[6] <- "lon"
gun_vio$date <- as.character(gun_vio$date)
data_new <- gun_vio %>%
  filter(state=="California")
data_new["year"]=str_sub( as.character(data_new$date),1,4)
data_new <- data_new %>%
  filter(year=="2017")
```

```{r echo=FALSE, fig.height=7, fig.width=5, warning=FALSE,message=FALSE}
qmplot(lon,lat,data=data_new,colour="red",zoom=9,size = I(1.1),maptype="toner-lite",main = "Gun violence in California in 2017")+
  theme(legend.position = "none")
```

As we could see from this plot, the gun violence incident is clustered at cities and roads, which means that gun violence incidents mostly happen where people live. 

Then we separated the US into two parts by Democratic Party and Republican Party. Does the gun violence is related to the party? We used boxplot trying to answer this question. The boxplots are shown below: 

```{r include=FALSE}
gun_vio_n_adjust = merge(gun_vio,population, by.x=c("state", "year"), by.y=c("state", "year"))

gun_vio_n_adjust['sum_killed_avg']<-(as.numeric(as.character(gun_vio_n_adjust$n_killed))+ as.numeric(as.character(gun_vio_n_adjust$n_injured)) )/as.numeric(as.character(gun_vio_n_adjust$population))*100000
gun_vio_n_adjust<-select(gun_vio_n_adjust,1,2,3,4,5,12,14)

gun_vio_n_adjust<-select(gun_vio_n_adjust,-3)
gun_vio_n_adjust_group<-gun_vio_n_adjust%>%
  group_by(year,month,state)%>%
  summarize("sum_killed_avg"=sum(sum_killed_avg))
gun_vio_n_adjust_group<-as.data.frame(gun_vio_n_adjust_group)
gun_vio_n_adjust_group_mrg<-merge(gun_vio_n_adjust_group,states_party,by.x='state',by.y='State')

p <- plot_ly(gun_vio_n_adjust_group_mrg, y = ~sum_killed_avg, color = ~Parties, title="Gun Violence by Parties",colors = c('blue',  'red'),type = "box") %>%
  layout(title="Boxplot of people killed/injured per 100,000 person by Party")
```

```{r echo=FALSE, warning=FALSE}
p
```

In the above plot, each dot represents a state, y-axis represents number of casualties per 100k person of each state all year together. The boxplots give us very interesting information. There are many outliers in Democratic Party, and the range of Democratic Party states is much larger. However, the average number of casulties in Democratic Party states is lower than that in Republican Party. 

### Which kind of gun violence is the most common?

Then, we look at the type of each incident. We summed the same kind of incidents from 2014 to 2018 in every state, the plot is shown below:  

```{r include=FALSE}
character<-select(gun_character,-1)
character['Mass Shooting']<-str_detect(character$incident_characteristics,'Mass Shooting')
character['Shot_Wounded/Injured']<-str_detect(character$incident_characteristics,'Shot - Wounded/Injured')
character['Possession (gun(s) found during commission of other crimes)']<-str_detect(character$incident_characteristics,'found during commission of other crimes')
character['Possession of gun by felon or prohibited person']<-str_detect(character$incident_characteristics,'Possession of gun by felon or prohibited person')
character['Drug involvement']<-str_detect(character$incident_characteristics,'Drug involvement')
character['Shots Fired']<-str_detect(character$incident_characteristics,'Shots Fired')
character['Bar/club incident - in or around establishment']<-str_detect(character$incident_characteristics,'Bar/club incident - in or around establishment')
# character['Armed robbery with injury/death and/or evidence of DGU found']<-str_detect(character$incident_characteristics,'Armed robbery with injury/death and/or evidence of DGU found')
character['Shot - Dead (murder, accidental, suicide)']<-str_detect(character$incident_characteristics,'Shot - Dead')
character['Officer Involved Shooting']<-str_detect(character$incident_characteristics,'Officer Involved Shooting')
character['Kidnapping/abductions/hostage']<-str_detect(character$incident_characteristics,'Kidnapping/abductions/hostage')
character['Brandishing/flourishing/open carry/lost/found']<-str_detect(character$incident_characteristics,'Brandishing/flourishing/open carry/lost/found')
character['Murder/Suicide']<-str_detect(character$incident_characteristics,'Murder/Suicide')
character['Drug involvement']<-str_detect(character$incident_characteristics,'Drug involvement')
character['Drive-by (car to street, car to car)']<-str_detect(character$incident_characteristics,'(car to street, car to car)')
character['Defensive Use']<-str_detect(character$incident_characteristics,'Defensive Use')
character['Domestic Violence']<-str_detect(character$incident_characteristics,'Domestic Violence')
character['Home Invasion']<-str_detect(character$incident_characteristics,'Home Invasion')
character['Institution/Group/Business']<-str_detect(character$incident_characteristics,'Institution/Group/Business')
character['Concealed Carry License - Perpetrator']<-str_detect(character$incident_characteristics,'Concealed Carry License - Perpetrator')
character['School Shooting']<-str_detect(character$incident_characteristics,'School Shooting')
character['Car-jacking']<-str_detect(character$incident_characteristics,'Car-jacking')
character['Hate crime']<-str_detect(character$incident_characteristics,'Hate crime')
character['House party']<-str_detect(character$incident_characteristics,'House party')
character['Workplace shooting']<-str_detect(character$incident_characteristics,'Workplace shooting')
character['Gun(s) stolen from owner']<-str_detect(character$incident_characteristics,'Gun(s) stolen from owner')
character['Accidental/Negligent Discharge']<-str_detect(character$incident_characteristics,'Accidental/Negligent Discharge')
character['Accidental Shooting']<-str_detect(character$incident_characteristics,'Accidental Shooting')
character['ATF/LE Confiscation/Raid/Arrest']<-str_detect(character$incident_characteristics,'ATF/LE Confiscation/Raid/Arrest')
character['Non-Shooting Incident']<-str_detect(character$incident_characteristics,'Non-Shooting Incident')
character['Shots Fired - No Injuries']<-str_detect(character$incident_characteristics,'Shots Fired - No Injuries')
character['Murder']<-str_detect(character$incident_characteristics,'Murder')
character['Officer shot']<-str_detect(character$incident_characteristics,'Officer shot')
character['Officer killed']<-str_detect(character$incident_characteristics,'Officer killed')
character['Officer shot']<-str_detect(character$incident_characteristics,'Officer shot')
character['Child killed']<-str_detect(character$incident_characteristics,'Child killed')
character['Child injured']<-str_detect(character$incident_characteristics,'Child injured')


character['Police Targeted']<-str_detect(character$incident_characteristics,'Police Targeted')
character['Gang involvement']<-str_detect(character$incident_characteristics,'Gang involvement')
character['Unlawful']<-str_detect(character$incident_characteristics,'Unlawful')
character['Armed_robbery']<-str_detect(character$incident_characteristics,'Armed robbery')
character['Sex_crime']<-str_detect(character$incident_characteristics,'Sex crime')
character['Domestic_violence']<-str_detect(character$incident_characteristics,'Domestic Violence')
character['Road_Rage']<-str_detect(character$incident_characteristics,'Road rage')
character['Hate_crime']<-str_detect(character$incident_characteristics,'Hate crime')

character['Alert_Only']<-str_detect(character$incident_characteristics,'ALERT ONLY')
character['Terrorism']<-str_detect(character$incident_characteristics,'Terrorism')

character=select(character,-1)
character_sum<-colSums(character[,-1])
character_sum_df_tp<-as.data.frame(as.list(character_sum))
character_sum_df<-gather(character_sum_df_tp)
```


```{r echo=FALSE,fig.width=6}
character_sum_df%>%filter(value>1000)%>%
ggplot(aes(x= fct_reorder(key, value)   ,y=value,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  coord_flip()+
  ggtitle("Key characteristics")+
  xlab("")+
  theme(legend.position = "none")+
  ggtitle("Number of Incidents with Certain Key Characteristics")
```

Seen from the statistic chart, we know that the characteristic of a gun violence are described about its cause and the damage it brought as well as some information abouth the arm gun. We could find that there are many different characteristic of incidents. The most common incident characteristic is shot-wounded/injured, the second common characteristic is shot-murder. 

**Mass Shooting**  

One damage that we would like to know is about the trends of mass shooting across the countries through these years.  
```{r echo=FALSE}
mass_shooting<-gun_vio%>%
  filter( `Mass.Shooting`==TRUE  )%>%
  filter(!is.na(lon))
mass_shooting$lon<-as.numeric(as.character(mass_shooting$lon))
mass_shooting<-mass_shooting%>% filter(lon> -140)

# get us state map
us_states <- map_data("state")


 ggplot()+
    geom_polygon(data=us_states, aes(x=long,y=lat,group=group),fill="#FFEDA0",color="grey")+
    geom_point(data=mass_shooting,aes(x=lon,y=lat,color=year),size=0.5)+
   ggtitle("Mass Shooting in the US from 2014 to March, 2018")
   
#heat_map_plotter("Florida",2017)
```

As shown in the map, we could find that this kind of incidents in different years are clustered in the countries. In California, most mass shootings occur in the area of San Fransisco and Los Angeles while other cities barely has one. And on the southern and esatern parts of the state, mass shootings are frequently seen there. And these states are also having higer rate of casualties in gun violence.  

**NUmber of Incidents that Children or Officers Killed or Injured in Gun Violence**   
From plot above, we found that the relative number of people killed to people injured and we would like to explore if that would be similar to different groups of people, officers and children are the groups of most concern.  

```{r echo=FALSE}
character_result<-subset(character_sum_df_tp, select=c('Officer.shot','Officer.killed','Child.injured', 'Child.killed'))

colnames(character_result)<-c('Officer injured','Officer killed','Child injured', 'Child killed')


character_result_ga<-gather(character_result)
character_result_ga%>%filter(value>=0)%>%
ggplot(aes(x= key ,y=value,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  coord_flip()+
  ggtitle("Number of incidents involving Officers or Children")+
  xlab("Key Cause Characteristics")+
  theme(legend.position = "none")
```

We found that though these two groups are having more injured when involved in gun violence incidents, it seems like children are more inclined to be in danger of death, so we could know that gun is a bigger threat to children.

**Major Causes of Gun Violence**  
We are now concerned about the the causes that brought about the gun violence. We are more interested about the intentional shooting as these would put more threat to the safety.

```{r echo=FALSE}
character_reason<-subset(character_sum_df_tp, select=c("Murder",'Home.Invasion','Terrorism','ATF.LE.Confiscation.Raid.Arrest', 'Police.Targeted','Gang.involvement','Car.jacking','Kidnapping.abductions.hostage','Armed_robbery','Sex_crime','Domestic.Violence','Road_Rage','Hate.crime'))
character_reason_ga<-gather(character_reason)
character_reason_ga%>%filter(value>=0)%>%
ggplot(aes(x= fct_reorder(key, value)   ,y=value,fill="#b30000"))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values=c("#b30000"))+
  coord_flip()+
  ggtitle("Number of Key Conflicts that Cause Gun Violence")+
  xlab("Key Cause Characteristics")+
  theme(legend.position = "none")

```

From the plot above, we could find that the most common conflicts that involved gun is the armed robbery. Then, incidents of ATF.LE Arrest and so on are ranking second, this is about the police carrying out duties, and it makes sense that it would arase many gun violences. Surprisingly, conflicts involving gang are fewer than domestic violence. It seems like the gun violence distributions follow the pattern taht the milder, the more frequently it would happen.


```{r echo=FALSE}
# character_year<-gun_character
year_df<-as.data.frame(as.Date(gun_character$date))
accidental_shooting<-character$`Accidental Shooting`
year_def<-cbind(year_df,accidental_shooting)
colnames(year_def)<-c("date","accidental_shooting")
year_def["year"]<-str_sub(as.character(year_def$date),1,4)
year_ded_group<-year_def%>%
  group_by(year)%>%
  summarize('accidental_shooting_sum'=sum(accidental_shooting),"incident_num"=n(),"non_accidental_shooting_sum"=n()-sum(accidental_shooting))
year_ded_group<-as.data.frame(year_ded_group)
year_ded_group_gather<-gather(year_ded_group,accidental_shooting_sum,non_accidental_shooting_sum,-year,-incident_num)

year_ded_group_gather["proportion"]<-year_ded_group_gather$non_accidental_shooting_sum/year_ded_group_gather$incident_num

colnames(year_ded_group_gather)<-c("year","incident_num","type","n","proportion")
year_ded_group_gather%>%
  filter(as.numeric(as.character( year_ded_group_gather$year))>2013)%>%
ggplot(aes(fill=type, y=proportion, x=year)) + 
  coord_flip()+
  geom_bar(position="fill", stat="identity")+
  ggtitle("Proportion of Acciendental Shooting from 2014 to 2018")
```

Apart from the threat that gun would bring, we would like to know more about the positive effect of the gun. What most people promotes the right of keeping are saying that the gun would help them defend themselves when their life are in danger. From the plot about the number of incidents that have certain characteristics, we can see that the number of incidents that have self defense is close to the number that have accidental shootings. Which means that the role that guns to defend a person might not be as effective as what people promote.

### When does gun violence happen?

Then we would like to explore the data in time series. Are the gun violence incidents more likely to happen in sepcific time? we plotted each gun violence incident by time from 2014 to 2018, 

```{r include=FALSE}
gun_vio$n_killed<-as.numeric(as.character(gun_vio$n_killed))
gun_vio$n_injured<-as.numeric(as.character(gun_vio$n_injured))
gun_vio<-gun_vio%>%
  filter(!is.na(n_killed) | !is.na(n_injured))
gun_vio_day<-gun_vio%>%
  group_by(date,year)%>%
  summarize('incident_number'=n(),'sum_killed'=sum(n_killed),'sum_injured'=sum(n_injured))
gun_vio_day <- gun_vio_day %>%
  filter(!is.na(sum_killed) | !is.na(sum_injured))
gun_vio_day<-as.data.frame(gun_vio_day)
gun_vio_day$date<-as.character(gun_vio_day$date)
gun_vio_day["month"]=str_sub(gun_vio_day$date,6,7)
gun_vio_day$date<-as.Date(gun_vio_day$date)
gun_vio_day_new<-gun_vio_day%>%
  group_by(year,month)%>%
  summarize('incident_number'=sum(incident_number),'sum_killed'=sum(sum_killed),'sum_injured'=sum(sum_injured))
gun_vio_day_new["adjusted_date"]<-as.Date( str_c(gun_vio_day_new$year,"-",gun_vio_day_new$month,"-15")  )

p <- plot_ly(gun_vio_day_new, x = ~adjusted_date, y = ~incident_number, name = 'incident_number', type = 'scatter', mode = 'lines+markers') %>%
  add_trace(y = ~sum_killed, name = 'sum_killed', mode = 'lines+markers') %>%
  add_trace(y = ~sum_injured, name = 'sum_injured', mode = 'lines+markers')%>%
  layout(title = 'Time Series of incident number, n_killed, n_injured',
         yaxis = list(title='number'),
         xaxis = list(title='time'))
```

```{r,fig.width=8,fig.width=8,echo=FALSE}
p
```

The above plot shows the sum of incident number, number of killed and number of injured of each month from 2014 to 2018. We could find that there are some peaks in this plot. We are very surprised to find that peaks are always in summer. We guessed that it might because people are easy to lose their mind when it's hot.  
Since some incidents don't involve any casualty, the numbers of incidents are higher than the sum of number killed (n_killed) and the number injured (n_injured). 

## Effects of Gun Possession

Since the numbers of firearm possessions are different among states, it might be related to the gun violence. 

### How does the gun possession influence the gun violence status?

Intuitively gun possession of each states may be related with gun violence. **total** in each rows in firearm background checking represents the number of transfer of firearms. It shows the gun possession of each state. We extracted the background check and gun violence records in 2017.

```{r include=FALSE}
back_check_sub <- back_check %>%
  select(1,2,27,29)
back_check_sub_2017 <- back_check_sub %>%
  filter(year=="2017")
back_check_gp_2017<-back_check_sub_2017%>%group_by(state,year)%>%
  summarize("casualty"=sum(totals))
back_check_gp_2017<-as.data.frame(back_check_gp_2017)



back_check_2017 <- merge(gun_vio_n_2017,back_check_gp_2017,by.x = "state",by.y = "state")
colnames(back_check_2017)<-c("state","X","year","casualty","year2","Background_Check")
population_sub <- population %>%
  filter(year=="2017")
back_check_2017 <- merge(back_check_2017,population_sub,by.x = "state",by.y = "state")
back_check_2017$Background_Check_avg<-back_check_2017$Background_Check/back_check_2017$population
back_check_2017$casualty_avg<-back_check_2017$casualty/back_check_2017$population
back_check_plot <- plot_ly(data = back_check_2017, x = ~Background_Check_avg, y = ~casualty_avg,
        marker = list(size = 6,
                       color = '#b30000',
                       line = list(color = '#b30000',
                                   width = 0.2))) %>%
  layout(title = 'Background Checking Number-Incident Number-2017',
         yaxis = list(zeroline = FALSE),
         xaxis = list(zeroline = FALSE))
```

```{r echo=FALSE, warning=FALSE,message=FALSE}
back_check_plot
```

From the plot above, we could find that most points are at the left bottom corner of the plot. There are two outliers in the plot. These two variables are slightly positively correlated. 

## Effects of GDP Data

We also would like to find whether there is something to do with the ecnomic situation. We choosed to use GDP of each state to represent the economic situation. Since there are few differences from 2014 to 2017, we choose to show the data in 2017. We extracted the GDP of the United States of year 2017 and the incident number of each state in 2017. In the scatterplot, each dot is a state, x-axis is GDP of this state, y-axis is incident of this state. 

```{r include=FALSE}
GDP_2017 <- GDP_sub %>%
  select(1,23)
gun_vio_n_2017 <- gun_vio_n %>%
  filter(year=="2017")
GDP_2017 <- merge(GDP_2017,gun_vio_n_2017,by.x = "GeoName",by.y = "state")

population_sub <- population %>%
  filter(year=="2017")
GDP_2017 <- merge(GDP_2017,population_sub,by.x = "GeoName",by.y = "state")
GDP_2017$GDP_avg<-GDP_2017$X2017/GDP_2017$population
GDP_2017$casualty_avg<-GDP_2017$total_TOLL/GDP_2017$population
GDP_2017_plot <- plot_ly(data = GDP_2017, x = ~GDP_avg, y = ~casualty_avg,
        marker = list(size = 6,
                       color = '#b30000',
                       line = list(color = '#b30000',
                                   width = 1))) %>%
  layout(title = 'GDP-Casualty-2017',
         yaxis = list(zeroline = FALSE),
         xaxis = list(title="GDP"))
```

```{r echo=FALSE, warning=FALSE,message=FALSE}
GDP_2017_plot
```


From the plot above, we could find that there is an outlier an the right up corner of the plot. Most of the points are at the left nottom of the plot. There are no relations between these two variables. 

## Combining these Variables

Now, we introduced gun violence data, background check data as well as GDP data. How do these variabled combine with each other? We use parallel coordinates charts to show their relations.

```{r include=FALSE}
Parcoord<-function(year)
{
  year<-as.character(year)
  back_check_year<-back_check[ which(back_check$year==year), ]
  back_check_state<-back_check_year%>%
    group_by(year,state)%>%
    summarize(total_year=sum(totals))
  
  gun_vio$date<-as.Date(gun_vio$date)
  gun_vio_tp<-gun_vio[which( format(as.Date(gun_vio$date, format="%d/%m/%Y"),"%Y")==year), ]
  gun_vio_tp<-subset(gun_vio_tp,select=c("date","state","n_killed","n_injured"))
  gun_vio_tp["n_sum"]<-as.numeric(as.character(gun_vio_tp$n_killed))+as.numeric(as.character(gun_vio_tp$n_injured))
  gun_vio_state<-gun_vio_tp%>%
    group_by(state)%>%
    summarize(n=sum(n_sum))

  check_vio<-merge(back_check_state, gun_vio_state, by.x="state", by.y="state")
  
  # uem_month$State<-str_trim(uem_month$State)
  # uem_rate<-uem_month[which(uem_month$Year==year), ]
  # uem_rate$State<-str_trim(uem_rate$State)
  
  GDP_year<-subset(GDP,select=c("GeoName",str_c("X",year)))
  
  colnames(GDP_year)<-c("state","GDP_total")
  population$year<-as.character(population$year)
  population_sub<-population[which(population$year==year),]
  GDP_avg=merge(GDP_year,population_sub,by.x="state",by.y="state")
  GDP_avg["GDP_avg"]=GDP_avg$GDP_total*1000000/GDP_avg$population
    
  check_vio_uem<-merge(check_vio, GDP_avg, by.x="state", by.y="state")
  
  check_vio_uem$n<-as.numeric(as.character( check_vio_uem$n))
  check_vio_uem$total_year<-as.numeric(as.character( check_vio_uem$total_year))
  check_vio_uem["checking_totals_avg"]<-check_vio_uem$total_year/check_vio_uem$population
  check_vio_uem["gun_violence_avg"]<-check_vio_uem$n/check_vio_uem$population
  
                              
  colnames(check_vio_uem)<-c("state","year","checking_totals","gun_violence_TOLL","GDP_total","Year", "population",  "GDP_avg","checking_totals_avg","gun_violence_avg")
  
  check_vio_uem<-check_vio_uem%>%
    filter(!is.na(gun_violence_TOLL))
  parcoords(check_vio_uem[c("state",  "checking_totals_avg","gun_violence_avg","GDP_avg")],rownames = FALSE,reorder = TRUE, brushMode="1D",alpha = .3, color = list(colorScale = "scaleOrdinal",colorBy = "state", colorScheme = "schemeCategory10"), withD3 = TRUE)
}
```

```{r}
Parcoord("2017")
```

Actually, patterns among those years are similar, so we only include the plot in 2017. In the plot above, most of the lines are at the bottom. 

Most lines between checking_totals_avg and gun_violence_avg are parellel, but there are still some crosses in between. We can conclude that the gun possession and gun violence are slightly positively correlated, which is consistent with the scatterplot above.  

There are lots of crosses between gun_viokence_avg and GDP_avg, so there is no clear relationship between gun violence and GDP.